---
title: "distribution-functions"
output: html_document
knit: (function(inputFile, encoding,...) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "~/Dropbox/featureSelection/tidyinfostats") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
#library(tidyinfostats)
# library(reticulate)
devtools::load_all("..")
```

## Distribution Functions

Dont use this unless you have very specific use cases.

The most sensible option is to use Distr6 <https://github.com/alan-turing-institute/distr6> for anything but the use case of testing various combined functions

```{r}
n = NormalDistribution$new(mean=0,sd=1)
n$plot(-5,5)
```
```{r}
ggplot(n$sample(100000),aes(x=x))+geom_histogram(bins = 500)
```

### Uniform density

```{r pressure, echo=FALSE}
n2 = UniformDistribution$new(min=2,max=4)
n2$plot(xmin=0,xmax=5)
```

### Log Normal distribution

Log normal plots are specified using a mode and a measure of variance

<!-- ```{python results='asis'} -->
<!-- from sympy import * -->
<!-- import numpy as np -->
<!-- x, y, z, t = symbols('x y z t') -->
<!-- k, m, n = symbols('k m n', integer=True) -->
<!-- f, g, h = symbols('f g h', cls=Function) -->
<!-- print('<math>'+mathml(integrate(1/x, x))+'</math>') -->
<!-- ``` -->

$$\mathit{mode}={{e}^{\mu-{{\sigma}^{2}}}} \\
\mathit{var}={{e}^{2\cdot \mu-{{\sigma}^{2}}}}\cdot \left( {{e}^{{{\sigma}^{2}}}}-1\right) $$


This can be rearranged to the following:

$$\mathit{var}+{{e}^{\mu}}\cdot \mathit{mode}-{{e}^{2\cdot \mu}}=0$$

$$ {e}^\mu = \frac{mode \pm \sqrt{mode^2 + 4var}}{2} $$
$$ \mu = log \left( \frac{mode \pm \sqrt{mode^2 + 4var}}{2} \right) $$

Which can be numerically solved for $\mu$ given the mode and variance, leaving the substitution for $\sigma$
$$\sigma=\sqrt{\mu-\mathrm{log}\left( \mathit{mode}\right) }$$

Which allows us to specify a log normal distribution with meaningful shape parameters.

```{r}
devtools::load_all("..")
n3 = LogNormalDistribution$new(mode=6,sd=2)
n3$plot(0,20)
n3$q(seq(0,1,length.out = 5))
```

```{r}
devtools::load_all("..")
n3 = LogNormalDistribution$new(mean=6,sd=2)
n3$plot(0,20)
```

```{r}
devtools::load_all("..")
n4 = KumaraswamyDistribution$new(mode=0.4,iqr=0.3)
n4$plot(0,1)
n5 = MirroredKumaraswamyDistribution$new(mode=0.6,iqr=0.3)
n5$plot(0,1)
```

The theoretical entropy value of this distribution is 

```{r}
paste0("Entropy: ",n3$theoreticalEntropy())
```

## Combining distributions

Combining two distributions is also possible, and the default labelling can be overridden

```{r}
cd = ConditionalDistribution$new()
cd$withDistribution(1,n)
cd$withDistribution(1,n2,"class 2")
cd$plot(xmin=-5,xmax=5)
```

```{r}
cd2 = ConditionalDistribution$new()
cd2$withDistribution(1,n4,"class 1")
cd2$withDistribution(1,n5,"class 2")
cd2$plot(xmin=0,xmax=1)
```

And we can sample from this combined distribution which is returned as a dataframe containing the class in "y"

```{r}
ggplot(cd$sample(10000),aes(x=x,fill=y))+geom_histogram(position="dodge",bins=40)
```

A limited number of theoretical values are available for the combined distribution (well numerically calculated)

```{r}
paste0("Mean: ",cd$theoreticalMean())
paste0("Variance: ",cd$theoreticalVariance())
paste0("Mutual information between x and y: ",cd$theoreticalMI())

```

For testing purposes it is also useful to be able to generate a sample from a random collection of distributions with approximately sensible defaults.
This is possible with a random distribution selector which will randomly pick from the known distributions

```{r}
cd3 = ConditionalDistribution$new()
cd3$withRandomDistributions(3)
cd3$plot(-5,5)
```

```{r}
# devtools::load_all("..")
testData = bloodResultsSimulation(1000)$data

#### Continuous probability estimation ----

ggplot(
  testData %>% group_by(feature,outcome) %>% probabilitiesFromContinuous(value, method="SGolay"),
  aes(x=value,y=p_x, colour=outcome)) + geom_point() + facet_wrap(vars(feature))

# debug(probabilitiesFromContinuous_SGolay)
# debug(applySGolayFilter)

ggplot(
  testDataLazy %>% group_by(feature,outcome) %>% probabilitiesFromContinuous(value, method="Kernel"),
  aes(x=value,y=p_x, colour=outcome)) + geom_point() + facet_wrap(vars(feature))

```

```