---
title: "Test estimate accuracy"
output: html_document
knit: (function(inputFile, encoding,...) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "~/Dropbox/featureSelection/tidyinfostats") })
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE}
library(tidyverse)
library(ggplot2)
library(standardPrintOutput)
library(tidyinfostats)
#devtools::load_all("..")
set.seed(101)
theme_set(standardPrintOutput::defaultFigureLayout())
```

## Define test distributions

### gaussians
gaussian test params
y0 = [ .4 .5 .8 ];          % the center of the gaussian
sigma_y = [ .2 .3 .25 ];    % the gaussian decay constant
p{2} = [ .2 1 0.5 ];        % the (normalized) amplitude p(x)

```{r}
gaussians = ConditionalDistribution$new()
gaussians$withDistribution(0.2,NormalDistribution$new(0.4,0.2))
gaussians$withDistribution(1,NormalDistribution$new(0.5,0.3))
gaussians$withDistribution(0.5,NormalDistribution$new(0.8,0.25))
```

### Uniform distributions
square wave test params

a = [ 0 .1 .2 ];            % the left side of each square wave
b_a = [ 1 1.1 1.1 ];        % the length in y of each square wave
p{1} = [ .2 1 0.5 ];        % the (normalized) amplitude p(x)

```{r}
squareWaves = ConditionalDistribution$new()
squareWaves$withDistribution(0.2,UniformDistribution$new(min=0,max=1))
squareWaves$withDistribution(1,UniformDistribution$new(min=0.1,max=1.2))
squareWaves$withDistribution(0.5,UniformDistribution$new(min=0.2,max=1.3))
```

### Log normals

```{r}
lognorm = ConditionalDistribution$new()
lognorm$withDistribution(0.5,LogNormalDistribution$new(mode=0.5,sd=0.25))
lognorm$withDistribution(1,LogNormalDistribution$new(mode=1.5,sd=1))
lognorm$withDistribution(0.3,LogNormalDistribution$new(mode=2.5,sd=0.5))
```

### plot our distributions

```{r}
gaussians$plot(-1,2)
standardPrintOutput::saveThirdPageFigure(filename="~/Dropbox/featureSelection/mutinfo/gaussiansDist")
```

```{r}
squareWaves$plot(-0.5,1.5)
standardPrintOutput::saveThirdPageFigure(filename="~/Dropbox/featureSelection/mutinfo/squareDist")
```

```{r}
lognorm$plot(0,5)
standardPrintOutput::saveThirdPageFigure(filename="~/Dropbox/featureSelection/mutinfo/lognormDist")
```


## Experiment 1

B. C. Ross, “Mutual information between discrete and continuous data sets,” PLoS One, vol. 9, no. 2, p. e87357, Feb. 2014 [Online]. Available: http://dx.doi.org/10.1371/journal.pone.0087357

The first experiment replicates Ross's findings around discretising data & KNN method for estimating MI but using 2 new algorithms and only the smaller sample size

```{r echo=FALSE}
#devtools::load_all("..")
experiment1 = function(distribution, sampleSize, reps) {
  
  set.seed(101)
  
  kout = NULL
  thMi = distribution$theoreticalMI()
  thMu = distribution$theoreticalMean()
  thSd = sqrt(distribution$theoreticalVariance())
  
  for (i in c(1:reps)) {
	  # i = 1
	  df = distribution$sample(sampleSize)
	  estMean = mean(df$x)
  	estSd = sd(df$x)
  	
  	# SGolay method
	  k = c(5,7,10,15,20,30,40)
	  kresult = sapply(k, function(k2) (calculateDiscreteContinuousMI(df, y, x, method = "SGolay", k_05=k2) %>% pull(I)))
  	kout = kout %>% bind_rows(tibble(
  			method = rep("SGolay",length(k)),
  			param = rep("filter width",length(k)),
  			test = rep(i,length(k)),
  			value = k,
  			estimated.MI = kresult,
  			estimated.Mean = rep(estMean,length(k)),
  			estimated.Sd = rep(estSd,length(k))
  	))
	
		# KWindow method
  	k = c(2:10)
		wresult = sapply(k, function(k2) (calculateDiscreteContinuousMI(df, y, x, method = "KWindow", k_05=k2) %>% pull(I)))
	  kout = kout %>% bind_rows(tibble(
  			method = rep("KWindow",length(k)),
  			param = rep("window width",length(k)),
  			test = rep(i,length(k)),
  			value = k,
  			estimated.MI = wresult,
  			estimated.Mean = rep(estMean,length(k)),
  			estimated.Sd = rep(estSd,length(k))
  	))
	  
	  # KNN method
  	k = c(2:10)
  	if (nrow(df)<500) {
		  w2result = sapply(k, function(k2) (calculateDiscreteContinuousMI(df, y, x, method = "KNN", k_05=k2) %>% pull(I)))
  	} else {
		  w2result = rep(NA,length(k))
		}
	  kout = kout %>% bind_rows(tibble(
  			method = rep("KNN",length(k)),
  			param = rep("knn distance",length(k)),
  			test = rep(i,length(k)),
  			value = k,
  			estimated.MI = wresult,
  			estimated.Mean = rep(estMean,length(k)),
  			estimated.Sd = rep(estSd,length(k))
  	))
	  
	  
	  # Discretise method
	  # binPow = seq(0.1,0.5,0.05)
	  # k = floor(sampleSize^binPow)
	  k = c(4,8,16,32,64,128,256)
	  nresult = sapply(k, function(k2) (calculateDiscreteContinuousMI(df, y, x, method = "DiscretiseByRank", discreteMethod="Histogram", bins=k2) %>% pull(I)))
	  kout = kout %>% bind_rows(tibble(
  			method = rep("Histogram by rank",length(k)),
  			param = rep("number bins",length(k)),
  			test = rep(i,length(k)),
  			value = k, #sampleSize/k,
  			estimated.MI = nresult,
  			estimated.Mean = rep(estMean,length(k)),
  			estimated.Sd = rep(estSd,length(k))
  	))
	  
	  # DiscretiseValue method
	  k = c(4,8,16,32,64,128,256)
	  nresult = sapply(k, function(k2) (calculateDiscreteContinuousMI(df, y, x, method = "DiscretiseByValue", discreteMethod="Histogram", bins=k2) %>% pull(I)))
	  kout = kout %>% bind_rows(tibble(
  			method = rep("Histogram by value",length(k)),
  			param = rep("number bins",length(k)),
  			test = rep(i,length(k)),
  			value = k,
  			estimated.MI = nresult,
  			estimated.Mean = rep(estMean,length(k)),
  			estimated.Sd = rep(estSd,length(k))
  	))
	  
	  # DiscretiseValue method
	  k = c(4,8,16,32,64,128,256)
	  nresult = sapply(k, function(k2) (calculateDiscreteContinuousMI(df, y, x, method = "DiscretiseByValue", discreteMethod="MontgomerySmith", bins=k2) %>% pull(I)))
	  kout = kout %>% bind_rows(tibble(
  			method = rep("Montgomery Smith",length(k)),
  			param = rep("number bins",length(k)),
  			test = rep(i,length(k)),
  			value = k,
  			estimated.MI = nresult,
  			estimated.Mean = rep(estMean,length(k)),
  			estimated.Sd = rep(estSd,length(k))
  	))
	  
	  # Compression method
	  k = c(1,2)
	  nresult = calculateDiscreteContinuousMI(df, y, x, method = "Compression") %>% pull(I)
	  kout = kout %>% bind_rows(tibble(
  			method = rep("Compress",length(k)),
  			param = rep("no param",length(k)),
  			test = rep(i,length(k)),
  			value = sampleSize/k,
  			estimated.MI = rep(nresult,length(k)),
  			estimated.Mean = rep(estMean,length(k)),
  			estimated.Sd = rep(estSd,length(k))
  	))
	  
  }
  
  kout = kout %>% mutate(
    theoretical.MI = thMi,
  	theoretical.Mean = thMu,
  	theoretical.Sd = thSd
  )
  
  
  
  return(kout)
}

exp1aData = experiment1(distribution=gaussians, sampleSize=400, reps=10)
exp1bData = experiment1(distribution=gaussians, sampleSize=10000, reps=10)
exp1cData = experiment1(distribution=squareWaves, sampleSize=400, reps=10)
exp1dData = experiment1(distribution=squareWaves, sampleSize=10000, reps=10)
exp1eData = experiment1(distribution=lognorm, sampleSize=400, reps=10)
exp1fData = experiment1(distribution=lognorm, sampleSize=10000, reps=10)

```

```{r}
# todo summarise experimental data
plotExperiment1 = function(df) {
  exp1aDataGrouped = df %>% group_by(method,param,value) %>% summarise( #,estimatedMean,estimatedSd,theoreticalMI,theoreticalMean,theoreticalSd) %>% summarise(
    estMIMean = mean(estimated.MI),
    estMIsd = sd(estimated.MI)
  )

  # p3 = 
  return(
    ggplot(exp1aDataGrouped, aes(x=value))+
      geom_hline(yintercept=min(df$theoretical.MI), colour="blue")+
  		geom_ribbon(aes(ymin=estMIMean-1.96*estMIsd,ymax=estMIMean+1.96*estMIsd), fill = "grey75",alpha=0.3)+
  		geom_line(aes(y=estMIMean))+ylab("estimated MI")+xlab("free parameter value")+
  		coord_cartesian(ylim = c(0,1))+
      #expand_limits(y=0)+
    	
      facet_wrap(vars(method),scales = "free_x")
  )
}
```

### Experiment 1: gaussians

Using a sample size of 400 and 100 repetitions of the test for bootstrapping

```{r}
plotExperiment1(exp1aData)
standardPrintOutput::saveHalfPageFigure(filename="~/Dropbox/featureSelection/mutinfo/gaussiansEstimated400")
```

Using a sample size of 10000 and 100 repetitions of the test for bootstrapping

```{r}
plotExperiment1(exp1bData)
standardPrintOutput::saveHalfPageFigure(filename="~/Dropbox/featureSelection/mutinfo/gaussiansEstimated10K")
```

### Experiment 1: uniform

Using a sample size of 400 and 100 repetitions of the test for bootstrapping

```{r cache=TRUE}
plotExperiment1(exp1cData)
standardPrintOutput::saveHalfPageFigure(filename="~/Dropbox/featureSelection/mutinfo/squareEstimated400")
```

Using a sample size of 10000 and 100 repetitions of the test for bootstrapping

```{r}
plotExperiment1(exp1dData)
standardPrintOutput::saveHalfPageFigure(filename="~/Dropbox/featureSelection/mutinfo/squareEstimated10K")
```

### Experiment 1: log normal

Using a sample size of 400 and 100 repetitions of the test for bootstrapping

```{r}
plotExperiment1(exp1eData)
standardPrintOutput::saveHalfPageFigure(filename="~/Dropbox/featureSelection/mutinfo/lognormEstimated400")
```

Using a sample size of 10000 and 100 repetitions of the test for bootstrapping

```{r}
plotExperiment1(exp1fData)
standardPrintOutput::saveHalfPageFigure(filename="~/Dropbox/featureSelection/mutinfo/lognormEstimated10K")
```

