---
title: "Test estimate accuracy"
output: html_document
knit: (function(inputFile, encoding,...) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "~/Dropbox/featureSelection/tidyinfostats") })
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE}
library(tidyverse)
library(ggplot2)
library(standardPrintOutput)
library(tidyinfostats)
#devtools::load_all("..")
set.seed(101)
theme_set(standardPrintOutput::defaultFigureLayout())
```

## Define test distributions

### gaussians
gaussian test params
y0 = [ .4 .5 .8 ];          % the center of the gaussian
sigma_y = [ .2 .3 .25 ];    % the gaussian decay constant
p{2} = [ .2 1 0.5 ];        % the (normalized) amplitude p(x)

```{r}
gaussians = ConditionalDistribution$new()
gaussians$withDistribution(0.2,NormalDistribution$new(0.4,0.2))
gaussians$withDistribution(1,NormalDistribution$new(0.5,0.3))
gaussians$withDistribution(0.5,NormalDistribution$new(0.8,0.25))
```

### Uniform distributions
square wave test params

a = [ 0 .1 .2 ];            % the left side of each square wave
b_a = [ 1 1.1 1.1 ];        % the length in y of each square wave
p{1} = [ .2 1 0.5 ];        % the (normalized) amplitude p(x)

```{r}
squareWaves = ConditionalDistribution$new()
squareWaves$withDistribution(0.2,UniformDistribution$new(min=0,max=1))
squareWaves$withDistribution(1,UniformDistribution$new(min=0.1,max=1.2))
squareWaves$withDistribution(0.5,UniformDistribution$new(min=0.2,max=1.3))
```

### Log normals

```{r}
lognorm = ConditionalDistribution$new()
lognorm$withDistribution(0.5,LogNormalDistribution$new(mode=0.5,sd=0.25))
lognorm$withDistribution(1,LogNormalDistribution$new(mode=1.5,sd=1))
lognorm$withDistribution(0.3,LogNormalDistribution$new(mode=2.5,sd=0.5))
```

# Experiment 3

Objectives:
* Look at sample size versus accuracy
** MI, Mean and SD with increasing sample size
** regression to predict SD of MI from est mean versus theoretical mean  & est versus theoretical SD

```{r}
#devtools::load_all("..")

experiment3 = function(distribution, reps, meth = c("KWindow","KNN","SGolay","DiscretiseByRank","DiscretiseByValue","Compression")) {
  
  set.seed(101)
  result = NULL
  
  thMi = distribution$theoreticalMI()
  thMu = distribution$theoreticalMean()
  thSd = sqrt(distribution$theoreticalVariance())
  
  j=0
  for (samples in c(8,16,32,64,128,256,512,1024,2048,4096,8192,16384)) {
  
    for (i in c(1:reps)) {
      
      j=j+1
      
      df = distribution$sample(samples)
      minGroupSize = min(df %>% group_by(y) %>% count() %>% pull(n))
      estMi = sapply(meth, function(m) {
        # browser()
        calculateDiscreteContinuousMI(df, y, x, method = m) %>% pull(I)
      })
      estMean = mean(df$x)
      estSd = sd(df$x)
    
      result = result %>% bind_rows(
        tibble(
          id = j,
          minGroupSize = minGroupSize,
          sample = samples,
          param = c(meth,"Mean","Std deviation"),
          theoretical = c(rep(thMi,length(estMi)),thMu,thSd),
          estimated = c(estMi,estMean,estSd)
        )
      )
    }
  }
  
  return(result)
}

# setup error

quantifyError = function(df) {
  return(
    df %>% mutate(
      absoluteError = estimated-theoretical,
      relativeError = ifelse(theoretical<0.05,NA,absoluteError/theoretical)
    ) %>% filter(!is.na(relativeError)) %>% group_by(sample,param) %>% summarise(
      minGroupSize =mean(minGroupSize),
      theoretical = mean(theoretical, na.rm=TRUE),
      rmse = sqrt(mean(absoluteError^2, na.rm=TRUE)),
      mae = mean(absoluteError, na.rm=TRUE),
      vae = var(absoluteError, na.rm=TRUE),
      nmae = mean(relativeError, na.rm=TRUE),
      nvae = var(relativeError, na.rm=TRUE),
      median_ae = quantile(absoluteError, probs=c(0.5), names=FALSE, na.rm=TRUE),
      upper_iqr_ae = quantile(absoluteError, probs=c(0.75), names=FALSE, na.rm=TRUE),
      lower_iqr_ae = quantile(absoluteError, probs=c(0.25), names=FALSE, na.rm=TRUE)
    )
  )
}



exp3aData = experiment3(lognorm,10)
exp3bData = experiment3(gaussians,10)
exp3cData = experiment3(squareWaves,10)
```

```{r}
# todo summarise experimental data
plotExperiment3 = function(df, components) {
  summary = df %>% group_by(sample,param) %>% quantifyError()
  return(ggplot(summary %>% filter(param %in% components),aes(x=sample))+
  		geom_line(aes(y=mae),colour="blue")+
  		geom_ribbon(aes(ymin=mae-1.96*vae,ymax=mae+1.96*vae),fill="blue",alpha=0.1)+
  		#geom_line(aes(y=estimatedMedian),colour="grey75")+
  		#geom_ribbon(aes(ymin=estimatedLower,ymax=estimatedUpper),fill="grey75",alpha=0.1)+
  		coord_cartesian(ylim = c(-0.25,0.25))+
      geom_hline(yintercept=0, colour="red")+
  		ylab("absolute error")+
  		xlab("sample size")+
    	facet_wrap(vars(param))+scale_x_log10())
}
```

### Log Normal

```{r}
# devtools::load_all("..")

plotExperiment3(exp3aData,c("Mean","Std deviation"))
plotExperiment3(exp3aData,c("KWindow","KNN","SGolay","DiscretiseByRank","DiscretiseByValue","Compression"))
standardPrintOutput::saveThirdPageFigure(filename="~/Dropbox/featureSelection/mutinfo/bootstrappingLogNorm")
```

### Gaussians

```{r}
#devtools::load_all("..")

plotExperiment3(exp3bData,c("Mean","Std deviation"))
plotExperiment3(exp3bData,c("KWindow","KNN","SGolay","DiscretiseByRank","DiscretiseByValue","Compression"))
standardPrintOutput::saveThirdPageFigure(filename="~/Dropbox/featureSelection/mutinfo/bootstrappingGaussians")
```

### Uniform

```{r}
# devtools::load_all("..")

plotExperiment3(exp3cData,c("Mean","Std deviation"))
plotExperiment3(exp3cData,c("KWindow","KNN","SGolay","DiscretiseByRank","DiscretiseByValue","Compression"))
standardPrintOutput::saveThirdPageFigure(filename="~/Dropbox/featureSelection/mutinfo/bootstrappingUniform")
```